---
title: "Homework 1"
output: github_document
editor_options: 
chunk_output_type: inline
---

**1. Consider the data set given below**

x <- c(0.18, -1.54, 0.42, 0.95)

And weights given by

w <- c(2, 1, 3, 1)


**Give the value  (a, b, c, or d) that minimizes the least squares equation **

**$\sum_{i=1}^N w_i(x_i - \mu)^2$**


a: 0.1471

b: 0.300

c: 0.0025

d: 1.077


Hmmm...how do we figure this out? Well, one way would be to just plug in values.
```{r}

#Let's just copy and paste in the R vectors from above.
x <- c(0.18, -1.54, 0.42, 0.95)
w <- c(2, 1, 3, 1)
#Create a vector of mu values
u <- c(0.1471,0.300,0.0025,1.077)

for(i in 1:4)
{
  newScore = sum(w*(x - u[i])^2)
  print(newScore)
}
```
**2. Consider the following data set**

x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)

y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

Fit the regression through the origin and get the slope treating y as the outcome and x as the regressor.

a: 0.59915

b: -0.04462

c: -1.713

d: 0.8263

```{r}
#here we try to explain y using x by estimating beta x, thus generating a line with slope beta x. Specifically in this case, we're going to force the line to go through the origin, (0,0). So force the line to go through (0,0). This is usually not a good solution unless we center our data. 
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)

y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

df = data.frame(x,y)

plot(y~x ,data = df,
     xlab = "X",
     ylab = "Y",
     main = "Y vs X",
     pch  = 20,
     cex  = 2,
     col  = "grey")
```


```{r}
Sxy = sum((x - mean(x)) * (y - mean(y)))
Sxx = sum((x - mean(x)) ^ 2)
Syy = sum((y - mean(y)) ^ 2)
c(Sxy, Sxx, Syy)
```
```{r}
beta_1_hat = Sxy / Sxx
beta_0_hat = mean(y) - beta_1_hat * mean(x)
c(beta_0_hat, beta_1_hat)
```

```{r}
model <- lm(y~x - 1)
model

model <- lm(y~x + 1)
model

model <- lm(y~x)
model

x_centered <- y-mean(y)
y_centered <- x-mean(x)
model <- lm(y_centered ~ x_centered)
model
```

```{r}
plot(y~x,data = df,
     xlab = "X",
     ylab = "Y",
     main = "Y vs X",
     pch  = 20,
     cex  = 2,
     col  = "grey")
abline(model, lwd = 3, col = "darkorange")
```

3. From the datasets package ‘mtcars’, fit the regression model with mpg as the outcome and weight (wt) and horsepower (hp) as the predictor. Use the matrix notation [insert latex] , give the slope coefficient and adjusted  as R output

```{r}
data("mtcars")
head(mtcars)
y <- mtcars$mpg
x1 <- mtcars$wt
x2 <- mtcars$hp

x0 <- rep(1,length(x1))
Y = as.matrix(y)
X = as.matrix(cbind(x0,x1,x2))
beta <- solve(t(X) %*% X) %*% (t(X) %*% Y)
beta

Y_hat <- X %*% beta
SST   = sum((Y - mean(Y)) ^ 2)
SSReg = sum((Y_hat - mean(Y)) ^ 2)

R2 = SSReg/SST
R2

# model <- lm(y~x1+x2)
# coef(model)
# summary(model)

```

4. Consider the following data set (used above as well).
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)
What is the intercept for fitting the model with x as the predictor and y as the outcome?

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

model <- lm(y~x)
summary(model)

```


5. You know that both the predictor and response have mean 0. What can be said about the intercept when you fit a linear regression?
a: It must be identically 0.
b: It is undefined as you have to divide by zero.

```{r}
#the constant (or Y intercept) will be the expected value of Y when predictor X is zero. The expected value of Y being the mean. 
y_zero <- y - mean(y)
x_zero <- x - mean(x)

model <- lm(y~x)
summary(model)

plot(y~x,data = df,
     xlab = "X",
     ylab = "Y",
     main = "Y vs X",
     pch  = 20,
     cex  = 2,
     col  = "grey",
     xlim= c(0,1))
abline(model, lwd = 3, col = "darkorange")
```


6. Consider the data given by
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
What value minimizes the sum of the squared distances between these points and itself?
a: 0.36
b: 0.573
c: 0.8
d: 0.44


```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
# calculate the mean
print(mean(x))
```
7. Let the slope having fit Y as the outcome and X as the predictor be denoted as $\sum_{i=1}^N w_i(x_i - \mu)^2$. Let the slope from fitting X as the outcome and Y as the predictor be denoted as . Suppose that you divide  by ; in other words consider .
What is this ratio always equal to?
a: Cor(Y,X)
b: Var(Y)/Var(X) 
c: 1
d: 2SD(Y)/SD(X)


