---
title: "diagnostics"
output: html_notebook
---
Note: incomplete

Load data and fit some models  
```{r}
data(women)
data(cars)
dir.data <- "/Users/jdstokes/repos/204b/data"
fpath <- file.path(dir.data,"alcoholarm.txt")
alcohol <- read.table(fpath,header = TRUE)
```

Generate some linear data
```{r}
#Generate x values from a uniform distribution
n <- 100
x <- runif(n, 0, 100)
beta_0 = 4
beta_1 = 3
sigma = 10

#Add noise
epsilon = rnorm(n, mean = 0, sd = sigma)
y = beta_0 + beta_1 * x + epsilon
linear = data.frame(y,x)

plot(x,y)
```


Generate some non-linear data
```{r}
#Generate x values from a uniform distribution
n <- 100
x <- runif(n, 0, 100)
beta_0 = 50
beta_1 = 30
sigma = 1
#Add noise
epsilon = rnorm(n, mean = 0, sd = sigma)
y = beta_0 + beta_1 * x^(-0.2) + epsilon
nonlinear = data.frame(y,x)

plot(x,y)
```


Modeling..
```{r}
#This data set gives the average heights and weights for American women aged 30–39.
fit_women   <- lm(weight ~ height, women)

#Speed of cars and distance it takes to stop
fit_cars    <- lm(dist ~ speed, cars)

#Is alcohol consumption related to muscle strength
fit_alcohol <- lm(strength ~ alcohol, alcohol)

# simulation
fit_linear <-  lm(y ~ x, linear)
fit_non <-  lm(y ~ x, nonlinear)
```


View four diagnostic plots
```{r}
plot(fit_linear)
```

Access Help page for lm() plot diagnostics
```{r}
?plot.lm
```

View all four plots as once
```{r}
par(mfrow=c(2,2))
plot(fit_linear)
```


Residuals vs Fitted  

This will let us see if there is a non-linear relationship between the predictor and outcome variables. What we'd like to see are equally spread residuals around a horizontal line. Below you'll see the residuals on the y-axis and fitted values on the x-axis. The residual = 0 line is plotted as the dotted line.

* Residuals should "bounce randomly" around the 0 line
* Residuals shold form a "horizontal band" around the 0 line
* If residuals that stand out from the random dot cloud, there may be outliers.

```{r}
plot(fit_linear, which = 1)
```

non-linear
```{r}
plot(fit_non, which = 1)
```

